---
title: "peakrefine"
output:   
    github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE,  fig.path = "README_figs/README-")
```

Refining peaks using metrics derived from Strand Cross-Correlation (SCC)

## Installation:
```{r install}
if(!require(devtools))
    install.packages("devtools")
devtools::install_github("jrboyd/peakrefine")
```

## Basic Usage
```{r use}
bam_file = "path/to/your/bam" #must be indexed
peaks_gr = rtracklayer::import("path/to/your/peaks")
corr_res = peakrefine::calcSCCMetrics(bam_file, peaks_gr, frag_sizes = 50:250)
```

Run this way, 100 peaks will take a bit under a minute.

Results will be automatically cached using BiocFileCache such that repeated 
calls with identical parameters load previous results.

## Running in parallel

peakrefine uses the mc.cores option to automatically split up peaks for
calculations.  If you run into memory or similar issues you should set n_splits
to a multiple of mc.cores (this applies to running in serial as well).  
The default is 1 job per core which may prove
too large.

```{r use_adv}
ncores = max(1, parallel::detectCores() - 1)
options(mc.cores = ncores)
peakrefine::calcSCCMetrics(bam_file, peaks_gr, frag_sizes = 50:250, n_splits = ncores * 3)
```

```{r use_real, echo=FALSE, eval=TRUE}
bam_file = "simulation/qc_framework/output_AF_RUNX1_ChIP/AF-MCF10A_RUNX1_pooled/AF-MCF10A_RUNX1_pooled.bam" #must be indexed
peaks_gr = rtracklayer::import("simulation/qc_framework/output_AF_RUNX1_ChIP/AF-MCF10A_RUNX1_pooled/AF-MCF10A_RUNX1_pooled_peaks.narrowPeak")
set.seed(1)
peaks_gr = sample(peaks_gr, 100)
options(mc.cores = 20)
corr_res = peakrefine::calcSCCMetrics(bam_file, peaks_gr, frag_sizes = 50:250)
```

## Using the output

```{r setup2, include=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```

Output is a named list:

```{r}
cbind(names(corr_res))
```

Getting read and fragment lengths:

```{r}
corr_res$read_length
corr_res$fragment_length
```

Other list items are data.tables:

```{r}
sapply(corr_res, class)
```

We'll need a couple more libraries here.

```{r}
library(data.table)
library(ggplot2)
```

```{r metric_corr}
metrics_dt = rbindlist(corr_res[3:5], use.names = TRUE, idcol = "metric")
theme_set(theme_classic())
ggplot(metrics_dt, aes(x = metric, y = correlation, color = metric)) + geom_boxplot()
```


Stable fragment correlation uses the calculated fragment size while flex always
uses the fragment size with the maximum SCC. Flex will report high correlation
from artifact peaks but is useful for assessing fragment size distribution.

```{r metric_shift}
ggplot(metrics_dt, aes(x = metric, y = shift, color = metric)) + geom_boxplot()
```


