---
title: "peakrefine"
output:   
    github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE,  fig.path = "README_figs/README-")
```

Refining peaks using metrics derived from Strand Cross-Correlation (SCC)

This package is under active development but the methods described here are
considered stable.  

Eventually peakrefine will be rolled into my existing bioconductor package, 
[seqsetvis](https://bioconductor.org/packages/release/bioc/html/seqsetvis.html), 
though still maintained separately here.

## Installation:
```{r install}
if(!require(devtools))
    install.packages("devtools")
devtools::install_github("jrboyd/peakrefine")
```

## Basic Usage
```{r use}
bam_file = "path/to/your/bam" #must be indexed
peaks_gr = rtracklayer::import("path/to/your/peaks")
corr_res = peakrefine::calcSCCMetrics(bam_file, peaks_gr, frag_sizes = 50:250)
```

Run this way, 100 peaks will take a bit under a minute.

Results will be automatically cached using BiocFileCache such that repeated 
calls with identical parameters load previous results.

## Running in parallel

peakrefine uses the mc.cores option to automatically split up peaks for
calculations.  If you run into memory or similar issues you should set n_splits
to a multiple of mc.cores (this applies to running in serial as well).  
The default is 1 job per core which may prove
too large.

```{r use_adv}
ncores = max(1, parallel::detectCores() - 1)
options(mc.cores = ncores)
peakrefine::calcSCCMetrics(bam_file, peaks_gr, 
                           frag_sizes = 50:250, n_splits = ncores * 3)
```

```{r use_real, echo=FALSE, eval=TRUE}
bam_file = "simulation/qc_framework/output_AF_RUNX1_ChIP/AF-MCF10A_RUNX1_pooled/AF-MCF10A_RUNX1_pooled.bam" #must be indexed
peaks_gr = rtracklayer::import("simulation/qc_framework/output_AF_RUNX1_ChIP/AF-MCF10A_RUNX1_pooled/AF-MCF10A_RUNX1_pooled_peaks.narrowPeak")
set.seed(1)
peaks_gr = sample(peaks_gr, 100)
options(mc.cores = 20)
corr_res = peakrefine::calcSCCMetrics(bam_file, 
                                      peaks_gr, 
                                      frag_sizes = 50:250)
```

## Using the output

```{r setup2, include=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```

Output is a named list:

```{r}
cbind(names(corr_res))
```

Getting read and fragment lengths:

```{r}
corr_res$read_length
corr_res$fragment_length
```

Other list items are data.tables:

```{r}
sapply(corr_res, class)
```

We'll need a couple more libraries here.

```{r}
library(data.table)
library(ggplot2)
```

There are 3 extracted correlation metrics:

* flex_fragment_correlation - the maximum correlation at any shift
* stable_fragment_correlation - the correlation for the average fragment size (determined by shift values in flex)
* read_correlation - the correlation for the read size (determined by mode of bam)

```{r metric_corr, fig.width=4, fig.height=3}
#items 3-5 of the results list contain these metrics
metrics_dt = rbindlist(corr_res[3:5], use.names = TRUE, idcol = "metric")
metrics_dt[, metric := sub("fragment", "frag", sub("_correlation", "", metric))]
theme_set(theme_classic())
ggplot(metrics_dt, aes(x = metric, y = correlation, color = metric)) + 
    geom_boxplot() + 
    guides(color = "none")
```


Stable fragment correlation uses the calculated fragment size while flex always
uses the fragment size with the maximum SCC. Flex will report high correlation
from artifact peaks but is useful for assessing fragment size distribution.

```{r metric_shift, fig.width=4, fig.height=3}
ggplot(metrics_dt, aes(x = metric, y = shift, color = metric)) + 
    geom_boxplot() + 
    guides(color = "none")
```

The last item in the list "full_correlation_results" stores the correlation 
for every value of shift.  The previous 3 are simply extracted from the 
full results for conveinence

```{r corr_profile, fig.width=4, fig.height=3}
to_plot = corr_res$full_correlation_results#[id %in% sample(unique(id), 3)]
to_plot$`Expected_Size` = "in expected range"
out_of_range_ids = unique(corr_res$flex_fragment_correlation[
    shift < 100 | shift > 200]$id)
to_plot[id %in% out_of_range_ids, `Expected_Size` := "out of range"]
ggplot(to_plot, aes(x = shift, y = correlation, group = id)) + 
    geom_path(size = 1.5, alpha = .1) + facet_wrap("Expected_Size") + 
    labs(title = "SCC profiles", 
         subtitle = "expected fragment size between 100-200 bp")
```


